Paper \#ZW10157

### 2019-01-31 paper/submission/theor-experiment.tex
adding Linda's edits to pg 23, end of tba section Added beam sizes at the pets location, reference to thesis

\begin{table}%[h!]
\begin{center}
  \caption{Convergence of benchmark problem with errors relative to
    hypervolume of sampled reference solution.}
  \label{tbl:bench_rms_error}
  \begin{tabular}{c|c|c}
    \hline\noalign{\smallskip}
    tot.\ function  & hypervolume & relative error\\
    evaluations    & & \\
    \noalign{\smallskip}\hline\noalign{\smallskip}
    100  &  0.859753 & $3.076 \times 10^{-1}$ \\
   % \noalign{\smallskip}\hline\noalign{\smallskip}
    200  &  0.784943 & $1.938 \times 10^{-1}$ \\
    500  &  0.685183 & $4.210 \times 10^{-2}$ \\
    900  &  0.661898 & $6.689 \times 10^{-3}$ \\
    1100 &  0.657615 & $1.749 \times 10^{-4}$ \\
    \noalign{\smallskip}\hline
  \end{tabular}
\end{center}
\end{table}

@@

\subsubsection{Hyper parameter Scan}
While the optimization problem and goals were well defined (Subsection \ref{awaproblem}), 
it was not clear what the best hyper parameters for the genetic algorithm would be.
These parameters include gene mutation probability, mutation probability, 
recombination probability, number of individuals, 
and number of generations to complete. 
Given the beam line in Fig.~\ref{awa-linac},
four small optimization experiments were done with various hyper parameters. 
Similar to the time step scan, 
the goal of this exercise was to determine which set of optimization
parameters strongly influence the results, 
and whether there was a time to solution difference.
From here on, we will reference each experiment as ex-1, ex-2, ex-3, and ex-4
as shown in Table \ref{extable}. 
\begin{table}%[h!]
	\begin{center}
		\caption{Input Parameters for initial twenty four hour AWA optimization experiments. 
			The gene mutation probability was equal to the mutation probability (not shown) in all four experiments. 
			The max number of individuals per generation was~80.}
		\label{extable}
		\begin{tabular*}{0.48\textwidth}{l ccc}
			& Gene Mutation Probability & Recombination Probability & Number of completed generations \\
			ex-1 &  0.1  & 0.9  &  96 \\
			ex-2 &  0.3  & 0.7  &  81 \\
			ex-3 &  0.8  & 0.2  &  53 \\
			ex-4 &  0.01 & 0.09 &  95 \\
		\end{tabular*}
	\end{center}
\end{table}

The maximum number of individuals per generation was fixed at 80. 
This number was chosen based on the node architecture, and the 
to prevent a prohibitive computational cost.  
Each experiment was allowed to run for twenty four hours, with 
a maximum generation limit of 100. 
We reduced the six objectives to four, 
and shortened the simulation time by moving the objectives further 
upstream to $s_1$ and $s_2$, the locations before and after the kicker, 
see Fig.~\ref{awa-tba}.

The objectives include: $\varepsilon_{x}\left(s = s_1\right)\text{, } \varepsilon_{x}\left(s = s_2\right)$, $\text{rms}_{s}\left(s = s_1\right)\text{, and }  \text{rms}_{s}\left(s = s_2\right)$. 
The OPAL input file for ex-3, and subsequent optimization runs in this paper, can be found at in following repository: 
\url{https://github.com/nneveu/awa-tba}.

@@

With these observations in mind, several beam parameters corresponding to
options on the Pareto Front in Fig.~\ref{fig:pareto1} were plotted and compared. 
A select result is shown in Fig.~\ref{fig:stat}.  
The maximum beam sizes are well below the beam pipe aperture limits as shown in Fig.~\ref{fig:stat}.
The solution is nearly round, which will increase changes of keeping the beam nearly round
as it travels to the last triplet in Fig.~\ref{awa-tba}.
Overall this solution is satisfactory, and meets all requirements for the new TBA beamline at the AWA.
The PETS (Power Extractor and Transfer Structure) is the physical device putting the tightest constraints on the beam parameters.  
At this location the aperture diameter is \SI{17.6}{mm}, and the desired bunch length is \SI{1.5}{m}.  
The optimization meets all requirements in general; and specifically at the PETS location, 
the beam parameters are $\sigma_x=\sigma_y=2.5$mm, with aa bunch length of XXXXX.

 Although the simulation results at the PETS location cannot yet be checked since the beamline is not yet completely installed, some checks have been possible for the partially completed beamline.  Simulations of the partial beamline have been done with good agreement to measurement.  For example, beam images were taken on the YAG screen located downstream of the kicker and the extracted beam sizes at different kicker angles compared to simulation.  Measurements showed an asymmetry between the x and y beam size that was also shown in simulation (see Fig. ).  This is likely due to the gun voltage since that had the most noticeable impact on the simulated asymmetry (the best match was at a gun voltage of 55 MV/m).  When it becomes possible to test the complete simulation solution, it is expected that there will not be perfect agreement with simulation; for example there may be errors in the fields of beamline elements.  However, the simulation can be a tool to track down the source of discrepancies, such as understanding the effect of the gun field on the beam symmetry.  More importantly, the optimization results should provide significant improvement in efficiency for achieving a good operational regime.

>>>>

\begin{table}%[h!]
\begin{center}
  \caption{Convergence of benchmark problem with errors relative to
    hypervolume of sampled reference solution.}
  \label{tbl:bench_rms_error}
  \begin{tabular}{c|c|c}
    \hline\noalign{\smallskip}
    tot.\ function  & hypervolume & relative error\\
    evaluations    & & \\
    \noalign{\smallskip}\hline\noalign{\smallskip}
    100  &  0.859753 & $3.076 \times 10^{-1}$ \\
    200  &  0.784943 & $1.938 \times 10^{-1}$ \\
    500  &  0.685183 & $4.210 \times 10^{-2}$ \\
    900  &  0.661898 & $6.689 \times 10^{-3}$ \\
    1100 &  0.657615 & $1.749 \times 10^{-4}$ \\
    \noalign{\smallskip}\hline
  \end{tabular}
\end{center}
\end{table}

@@

\subsubsection{Hyper parameter Scan}
While the optimization problem and goals were well defined (Subsection \ref{awaproblem}), 
it was not clear what the best hyper parameters for the genetic algorithm would be.
These parameters include gene mutation probability, mutation probability, 
recombination probability, number of individuals, 
and number of generations to complete. 
Given the beam line in Fig.~\ref{awa-linac},
four small optimization experiments were done with various hyper parameters. 
Similar to the time step scan, 
the goal of this exercise was to determine which set of optimization
parameters strongly influence the results, 
and whether there was a time to solution difference.
From here on, we will reference each experiment as ex-1, ex-2, ex-3, and ex-4
as shown in Table \ref{extable}. 
\begin{table}%[h!]
	\begin{center}
		\caption{Input Parameters for initial twenty four hour AWA optimization experiments. 
			The gene mutation probability was equal to the mutation probability (not shown) in all four experiments. 
			The max number of individuals per generation was~80.}
		\label{extable}
		\begin{tabular*}{0.48\textwidth}{l c|c|c}
			& Gene Mutation \qquad& Recombination \qquad& Number of completed\vspace{-0.75em}\\
			& Probability \qquad& Probability \qquad&  generations \\
			\midrule
			ex-1 &  0.1  & 0.9  &  96 \\
			ex-2 &  0.3  & 0.7  &  81 \\
			ex-3 &  0.8  & 0.2  &  53 \\
			ex-4 &  0.01 & 0.09 &  95 \\
		\end{tabular*}
	\end{center}
\end{table}

The maximum number of individuals per generation was fixed at 80. 
This number was chosen based on the node architecture, and the 
to prevent a prohibitive computational cost.  
Each experiment was allowed to run for twenty four hours, with 
a maximum generation limit of 100. 
We reduced the six objectives to four, 
and shortened the simulation time by moving the objectives further 
upstream to $s_1$ and $s_2$, the locations before and after the kicker, 
see Fig.~\ref{awa-tba}.  
The objectives include: $\varepsilon_{x}\left(s = s_1\right)\text{, } \varepsilon_{x}\left(s = s_2\right)$, $\text{rms}_{s}\left(s = s_1\right)\text{, and }  \text{rms}_{s}\left(s = s_2\right)$. 
The OPAL input file for these cases and all subsequent optimization runs in this paper, 
can be found at in following repository: \url{https://github.com/nneveu/awa-tba}.

@@

With these observations in mind, several beam parameters corresponding to
options on the Pareto Front in Fig.~\ref{fig:pareto1} were plotted and compared. 
A select result is shown in Fig.~\ref{fig:stat}.  
The maximum beam sizes are well below the beam pipe aperture limits as shown in Fig.~\ref{fig:stat}.
The solution is nearly round, which will increase changes of keeping the beam nearly round
as it travels to the last triplet in Fig.~\ref{awa-tba}.
Overall this solution is satisfactory, and meets all requirements for the new TBA beamline at the AWA.
A Power Extractor and Transfer Structure (PETS) will be located after the septum and quadrupoles in Fig.~\ref{awa-tba}.
This device is putting the tightest constraints on the beam parameters.  
The aperture diameter is \SI{17.6}{mm}, and bunch length should be below \SI{2}{mm}
to facilitate large power extraction~\cite{PETSeq}.

The optimization results shown here were used in part to meet all TBA requirements. 
After adjustments to include 3D rf field maps and CSR in the dipoles~\cite{neveuthesis}, 
the max beamsizes at the PETS location were below \SI{10}{mm} with a bunch length of $rms_z=\SI{1.6}{mm}$. 
Although the simulation results at the PETS location cannot yet be verified, since the beamline is not yet completely installed, 
some checks have been possible for the partially completed beamline.  
Simulations of the partial beamline have been done.
For example, beam images were taken on the YAG screen located downstream of the kicker and 
the extracted beam sizes at different kicker angles compared to simulation~\cite{neveuthesis}.  
When it becomes possible to test the complete simulation solution, it is expected that there will not be perfect agreement with simulation;
for example there may be errors in the fields of beamline elements.  However, the simulation can be a tool to track down the source of
discrepancies, such as understanding the effect of the gun field on the beam symmetry.  
More importantly, the optimization results should provide significant improvement in efficiency for achieving a good operational regime.
